---
title: "Conjoint with LLMs"
output: github_document
author: "Shimu Liu"
date: "2025-05-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Introduction:
In this workshop, we'll demonstrate how to simulate a conjoint analysis 
based on Brand et al.(2024)'s paper using LLMs via API in R. Here's the overall 
guideline:

1. Design product profiles (e.g., laptops)

2. Generate prompts for LLM

3. Send requests using API

4. Parse responses to simulate consumer choices

Start Implementation:

1.a Load Libraries
```{r}
# Load required packages
# httr and jsonlite are used for sending HTTP requests and handling JSON responses
library(httr)       # For API requests
library(purrr)      # For functional programming
library(dplyr)      # For data manipulation
library(tibble)     # For tidy data frames
library(jsonlite)   # For JSON handling
library(tidyr)      # For data tidying
library(readr)      # For reading/writing CSV files
```

1.b Set API Key
(https://openrouter.ai/)
OpenRouter is a unified API platform that provides streamlined access to a wide array of large language models (LLMs)

We define our OpenRouter API key and specify which LLM to use. For this example, we use the "deepseek/deepseek-chat-v3-0324:free" model. You can substitute any other OpenRouter-compatible model here.

```{r}
# Set your OpenRouter API key
#Sys.setenv(OPENROUTER_API_KEY = "Your API Key")

Sys.setenv(OPENROUTER_API_KEY =  "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")


# DeepSeek model specification
deepseek_model <- "deepseek/deepseek-chat-v3-0324:free"


```

2. Design Conjoint Profiles
```{r}
# Define product attributes
brands <- c("Microsoft Surface Laptop", "Apple Macbook Air")
prices <- c("$1000", "$1200", "$1400")
rams <- c("8GB", "16GB", "32GB")
storages <- c("256GB", "512GB", "1TB")
sizes <- c("13in", "15in")

# Create all possible product combinations
generate_product_profiles <- function() {
  expand.grid(
    Brand = brands,
    Price = prices,
    Size = sizes,
    RAM = rams,
    Storage = storages,
    stringsAsFactors = FALSE
  ) %>% 
    mutate(ProductID = paste0("P", row_number()))
}

products <- generate_product_profiles()
print(products)
```

3. Generate Prompts
```{r}
# Defines a function to generate comparison prompt
generate_prompt <- function(product1, product2) {
  #sprintf() is used to create a formatted string
  prompt1 <- sprintf("%s, Price: %s, RAM: %s, Storage: %s, Size: %s",
                    product1$Brand, 
                    product1$Price, 
                    product1$RAM, 
                    product1$Storage, 
                    product1$Size)
  
  prompt2 <- sprintf("%s, Price: %s, RAM: %s, Storage: %s, Size: %s",
                    product2$Brand, 
                    product2$Price, 
                    product2$RAM, 
                    product2$Storage, 
                    product2$Size)
  
  # Create the final prompt
  # 1. system message
  sys_msg <- list(
    role    = "system",
    content = paste(
      "You are a customer. You are selected at random while shopping",
      "for laptops to participate in a survey. The interviewer will describe the",
      "options you saw while shopping and ask you to report which option you chose",
      "to purchase. Whenever two options are shown, you can also choose a third",
      "option which is not to purchase anything that day."
    )
  )
  
  # 2. user message
  usr_msg <- list(
    role    = "user",
    content = sprintf(
      paste(
        "Option 1: %s\n",
        "Option 2: %s\n\n",
        "Which option do you choose? Respond ONLY with:\n",
        "- '1' for Option 1\n",
        "- '2' for Option 2\n",
        "- '0' if you choose not to purchase anything"
      ),
      prompt1, prompt2
    )
  )
  
  # 3. bundle them for the API call
  combined_content <- paste(sys_msg$content, usr_msg$content, sep = "\n\n")
  return(combined_content)
}


# Function to create all unique product pairs with prompts
generate_product_pairs <- function(products, iterations) {
  all_pairs <- combn(products$ProductID, 2, simplify = FALSE)
  sampled_pairs <- sample(all_pairs, min(iterations, length(all_pairs)))
  
  map_dfr(sampled_pairs, ~ {
    p1 <- products %>% filter(ProductID == .x[1])
    p2 <- products %>% filter(ProductID == .x[2])
    tibble(
      Product1_ID = p1$ProductID,
      Product2_ID = p2$ProductID,
      Prompt = generate_prompt(p1, p2)
    )
  })
}

pairs_df <- generate_product_pairs(products, iterations = 20)
#Inspect the first few rows of the generated prompts
head(pairs_df)
pairs_df$Prompt[[1]]

```

4. Send Requests to API
Check for more standard guidance:
https://openrouter.ai/deepseek/deepseek-chat-v3-0324:free/api
```{r}
query_llm <- function(prompt) {
  res <- POST(
    url = "https://openrouter.ai/api/v1/chat/completions",
    add_headers(
      "Authorization" = paste("Bearer", Sys.getenv("OPENROUTER_API_KEY")),
      "User-Agent" = "demo"
    ),
    encode = "json",
    body = toJSON(list(
      model = "deepseek/deepseek-chat-v3-0324:free",
      messages = list(list(role = "user", content = prompt)),
      max_tokens = 150
    ), auto_unbox = TRUE)
  )

  raw <- content(res, "text", encoding = "UTF-8")
  #print("=== Raw response ===")
  #print(raw)

  result <- fromJSON(raw, simplifyVector = FALSE)
  content_text <- result$choices[[1]]$message$content

  # Use regex to extract only the first 0, 1, or 2
  library(stringr)
  choice <- str_extract(content_text, "[012](?=\\s|$|\\b)")

  print(paste("Model replied (cleaned):", choice))
  return(choice)
}


```

5.a Collect Responses
```{r}

# Run the study (comment out during testing)
#slice the first 2 rows
pairs_test <- pairs_df |> slice_head(n = 5)

rate_limit_pause <- 10  
print(pairs_test) 

responses <- pairs_test %>%
  mutate(Response = map_chr(Prompt, ~ {
    Sys.sleep(rate_limit_pause)
    query_llm(.x)  # capture the choice directly
  }))

head(responses)
```

Reference:
Brand, James, et al. “Using LLMs for Market Research.” SSRN Electronic Journal, 2023, https://doi.org/10.2139/ssrn.4395751.
